---
title: "CUNY 607 - Final Project"
author: "Kishore Prasad"
date: "May 8, 2016"
output: html_document
params:
    proj_user: "sql5118904"
    proj_pwd: "wUQcwMzwUl"
    proj_db: "sql5118904"
    proj_host: "sql5.freemysqlhosting.net"
---

```{r}
#mydb = dbConnect(MySQL(), user=params$proj_user, password=params$proj_pwd, dbname=params$proj_db, host=params$proj_host)

proj_user <- "sql5118904"
proj_pwd <- "wUQcwMzwUl"
proj_db <- "sql5118904"
proj_host <- "sql5.freemysqlhosting.net"


```




#Exploratory Data Analysis

##Does the Month have an impact on IPO Result?

Seems like months in the second half of the year have a better chance of succeeding instead of the first half.

```{r results = 'asis', comment = NA}

# Courtesy : http://walkerke.github.io/2014/01/hispanic-pop-by-state/

data.analysis <- melt(prop.table(table(AllData$ListingMonth, AllData$IPOresult),1), variable.name = 'IPOresult', value.name = 'Share')

d1 <- dPlot(
    x = "Share", 
    y = "Var1", 
    groups = "Var2", 
    data = data.analysis, 
    type = 'bar')

d1$xAxis(type = "addPctAxis")
d1$yAxis(type = "addCategoryAxis", orderRule = "Var1")
d1$legend( x = 60, y = 10, width = 700, height = 20, horizontalAlign = "left", orderRule = "Var2")

d1$show("iframesrc", cdn = TRUE)

```


##Impact of Over / Under subscription
Next I would like to see if Subscription rate has any correlation with listing gains.

There seems to be some correlation between the two.

```{r}

ggplot(data=AllData, aes(x=AllData$SubscriptionRate, y=AllData$ListingGains)) + geom_point()
m3 <- lm(AllData$ListingGains ~ AllData$SubscriptionRate, data = AllData)
summary(m3)

plot(AllData$ListingGains ~ AllData$SubscriptionRate)
abline(m3)

```

##Impact of favorable Review score
Do favorable review score give better listing gains? 

I dont see any specific impact of the review score as of now.

```{r}
ggplot(data=AllData, aes(x=AllData$ReviewScore, y=AllData$ListingGains)) + geom_point()
m3 <- lm(AllData$ListingGains ~ AllData$ReviewScore, data = AllData)
summary(m3)

plot(AllData$ListingGains ~ AllData$ReviewScore)
abline(m3)

```

#Divide the data into Train and Test dataset

```{r}
train.dataset <- subset(AllData, ListingYear  %in% c(2007:2014))
test.dataset <- subset(AllData, ListingYear  %in% c(2015:2016))

train.dataset <- train.dataset[,-c(1,6,8)]
test.dataset <- test.dataset[,-c(1,6,8)]

model <- glm(IPOresult ~.,family=binomial,data=train.dataset)
summary(model)

```


#Logistic regression
Having done some exploratory analysis, I would like to now come up with a regression equation for Successful or Unsuccessful IPO. 

```{r}

library(randomForest)
fit <- randomForest(IPOresult ~ ., train.dataset,ntree=500)
summary(fit)
predicted= predict(fit,x_test)


m_ipores <- lm(IPOresult ~ ., data = AllData)
summary(m_ipores)

```


#Random Forests

##Training the model

##Testing the model

##Evaluating the model
###Contingency Table
###Gains chart / Lift Chart


#Appendix

The below section was used to gather the data.

```{r eval=FALSE}


```
