---
title: "Moving data from MySQL to Neo4j"
author: "Kishore Prasad"
date: "April 20, 2016"
output: html_document
---


In this Project, we will extract some data from the  MySQL database - flights and create a Neo4j database. We will use the following tables from flights:

Airlines
Airports
Flights (a sample of 10,000 rows)
planes

It is assumed that you have MySQL 

Here are the various node types we will create:

Airline
Orig
Dest
Flight
Plane
AircraftType

The below are the relations that we will create

Flight OWNED_BY Airline
Flight FLEW_FROM Orig
Flight FLEW_TO Dest
Flight BY_PLANE Plane
Plane OF_TyPE AircraftType


```{r}
library(RMySQL)
library(RNeo4j)

```

# Get data from MySQL

In this section we connect to MySQL and create data frames from all the tables in the database.

```{r}
mydb = dbConnect(MySQL(), user='root', password='root', dbname='flights', host='localhost')

airlines = dbGetQuery(conn = mydb, statement = "Select * from airlines")
planes = dbGetQuery(conn = mydb, statement = "Select * from planes")
aircraft_types =  dbGetQuery(conn = mydb, statement = "Select distinct model, manufacturer, type, engines, seats, speed from planes")

airports = dbGetQuery(conn = mydb, statement = "Select * from airports")
flights = dbGetQuery(conn = mydb, statement = "Select * from flights")
```

# Create data structure in Neo4j

In this section we connect to Neo4j and create the structure for the flights database. 

Note: In my installation, I am disabling authentication.

```{r}

graph = startGraph("http://localhost:7474/db/data/")
clear(graph, input = FALSE)

# create the nodes for Airlines
for (i in 1:nrow(airlines)){
    createNode(graph, "Airline", carrier = airlines[i,1], name = airlines[i,2])
}
addConstraint(graph, "Airline", "carrier")


summary(graph)
```

