---
title: "CUNY - DAM - Week 6 Assignment"
author: "Kishore Prasad"
date: "March 5, 2016"
output: html_document
---

Read the raw data file from github.

```{r}
library(RCurl)
library(tidyr)
library(dplyr)
library(stringr)

flt_data<-read.csv("https://raw.githubusercontent.com/kishkp/CUNY-DAM/master/AirlineStatus.csv", stringsAsFactors = FALSE, check.names = FALSE, na.strings = c("", "NA"))
```


We now start the task of tidying up the data.

The first task is to make Airline value repeat in the following rows. That is, we fill in values for the airline.

```{r}

flt_data <- fill(flt_data, Airline)

```

Next we remove all blank rows. To do this we check the Status column and remove all NA rows.

```{r}

flt_data <- filter(flt_data, !is.na(Status))
```

We now need to have the all the cities in its own column called "Cities"

```{r}

flt_data <- gather(flt_data, cities, flights, -c(1, 2), na.rm = TRUE)

```

The final transformation is to split status and have a column each for "delayed" and "ontime". Before we do that we want to remove all space between the words "on" and "time" and then proceed with the column split.

```{r}

flt_data[,2] <- unlist(str_replace_all(flt_data[,2], " ", ""))

flt_data <- spread(flt_data, key = Status, value= flights )

```

Now we are ready for analysis.

Lets calculate the ratio of delayed to ontime.

```{r}

flt_data <- mutate(flt_data, del_ot_ratio= delayed / ontime)

```

Using this ratio, we now calculate the average ratio to compare between the 2 airlines.

```{r}

flt_data %>% group_by(Airline) %>% summarise(mean(del_ot_ratio))

```

We find that Alaska has a better delayed to ontime ratio than Am West.

Next we find the standard deviation for the 2 airlines.

```{r}

flt_data %>% group_by(Airline) %>% summarise(sd(ontime))

```

Here we find that Alaska has less standard deviation than Am West. 

**After analysis, we can conclude that Alaska has a better performance record than Am west.**

