---
title: "Week 3 Assignment"
author: "Kishore Prasad"
date: "February 14, 2016"
output: html_document
---

In the first part of this Assignment we created the tb_cases.csv in MySQL and exported the .csv file to the local file system.

We then copied the csv file to a github repository. We now use this repository to complete the rest of the assignment in R

First we create a dataframe with the tb_cases. To do this we get the csv from github and assign column names to the data (the original csv file did not have the column names in the csv)

```{r}
library(RCurl)

datafile<-getURL("https://raw.githubusercontent.com/kishkp/CUNY-DAM/master/tb_cases.csv", ssl.verifypeer = FALSE )

tb_cases<-read.csv(text = datafile, header=FALSE, stringsAsFactors = FALSE)

colnames(tb_cases)<- c("country", "year","tb_cases")

head(tb_cases)
```

We now do the same for the population file as well. However in case of population, the headers exist.

```{r}
datafile<-getURL("https://raw.githubusercontent.com/kishkp/CUNY-DAM/master/population.csv", ssl.verifypeer = FALSE )

population<-read.csv(text = datafile, header=TRUE, stringsAsFactors = FALSE)

head(population)
```

We now merge both the above data frames on the common key of country and year. We do a full outer join so that we dont miss out on any of the rows in both the data frames.

```{r}

tb_rate<-merge(tb_cases,population, all=TRUE)

head(tb_rate)
```

We then create a calculated column for the TB rate in the population and add it to the newly created data frame.

```{r}
tb_rate <- cbind(tb_rate, rate = tb_rate$tb_cases / tb_rate$population)

head(tb_rate)
```

Finally we drop all other columns and retain only the required columns of country, year and rate.

```{r}
final_tb_rate_data <- tb_rate[,c("country", "year", "rate")]

head(final_tb_rate_data)
```
